---
title: "CHASE Generate test dataset"
author: "NIVA Denmark"
date: "28/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Method

For testing purposes, we used indicator data from the HOLAS II assessment. This data was available with indicators per station. The indicator data came from the file **CHASEinput060318.xlsx** sheet **By station (for BSII only)**.

Using shape files for level 3 assessment units and level 4 assessment units, we mapped the station positions to assessment units, creating two input data sets for the CHASE assessment:

1) [assessmentdata_holas_ii.csv](input/assessmentdata_holas_ii.csv)

2) [assessmentdata_holas_ii_L4.csv](input/assessmentdata_holas_ii_L4.csv)

### Shape files

The shapes files used were downloaded from the HELCOM meeting portal: https://portal.helcom.fi/meetings/CORESET%20II%202-2014%20joint/MeetingDocuments/Forms/Documents.aspx

The original shape files contain several columns, including GES status, polygon area, etc. For the purpose of these tests, we are only interested in the names of the assessment units. They were therefore modified using the following code, to reduce their size.

```{r shapefiles, eval=FALSE}

library(tidyverse)
library(sf)

units3 <- read_sf(dsn = "../gis/CORESET/CORESET Assessment maps", 
                  layer = "HELCOM_AssessmentUnits_Scale3")
units3 <- units3 %>%
  dplyr::select(Name,HELCOM_ID)

units4 <- read_sf(dsn = "../gis/CORESET/CORESET Assessment maps", 
                  layer = "HELCOM_AssessmentUnits_Scale4")
units4 <- units4 %>%
  dplyr::select(HELCOM_ID) %>%
  filter(!is.na(HELCOM_ID))

st_write(units3, "assessment_units/AssessmentUnits3.shp",append=F)
st_write(units4, "assessment_units/AssessmentUnits4.shp",append=F)

```



```{r indicators}
knitr::opts_chunk$set(echo = TRUE)
```


## R Markdown


```{r cars}
summary(cars)
```

```{r pressure, echo=FALSE}
plot(pressure)
```
